\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-02-08 doc/publications/plos16/plos16.tex
fix my ref

In a previous paper, we clustered individuals from the 1000 Genomes Project using k-means clustering \cite{Obrien2016}.
\marginpar{reference is missing}
Although we were able to cluster individuals on all 80 million variants, it was relatively slow, with the entire
process (pre-processing and clustering) taking over 30 hours.
We investigate building random forest models on this data. However, we observed this process to not only be slower
than building a k-means model, but does not scale to as many dimensions (see Table~\ref{timingtable}). To keep
the following comparisons fair, we limit the number of executors available to Spark. On the relatively low-dimensional
phase1\_chr22 dataset (1,092 x 490,036), we observed a runtime of under 8 minutes. Although slower than Ranger, 
presumably due to the overhead of Spark's distributed platform on this relatively small dataset, we observe Spark ML to
be about 16 minutes faster than R. Moving to the larger dataset, phase3\_chr1 (2,504 x 6,450,364), we observe an increase
in Spark ML's runtime to over 8 hours. Moving to our next largest dataset, phase3\_chr1\-3 (2,504 x 19,328,051), and the
job fails to complete within our arbitrary cutoff time of 20 hours.

>>>>

In a previous paper, we clustered individuals from the 1000 Genomes Project using k-means clustering \cite{OBrien2015}.
Although we were able to cluster individuals on all 80 million variants, it was relatively slow, with the entire
process (pre-processing and clustering) taking over 30 hours.
We investigate building random forest models on this data. However, we observed this process to not only be slower
than building a k-means model, but does not scale to as many dimensions (see Table~\ref{timingtable}). To keep
the following comparisons fair, we limit the number of executors available to Spark. On the relatively low-dimensional
phase1\_chr22 dataset (1,092 x 490,036), we observed a runtime of under 8 minutes. Although slower than Ranger, 
presumably due to the overhead of Spark's distributed platform on this relatively small dataset, we observe Spark ML to
be about 16 minutes faster than R. Moving to the larger dataset, phase3\_chr1 (2,504 x 6,450,364), we observe an increase
in Spark ML's runtime to over 8 hours. Moving to our next largest dataset, phase3\_chr1\-3 (2,504 x 19,328,051), and the
job fails to complete within our arbitrary cutoff time of 20 hours.
